---
title: "Who, What, Where, Why, How? Sports"
author: "Brady Monks"
date: "12/7/2020"
output: 
  html_document:
    code_folding: hide
---

I wanted to decipher where professional athletes come from. I started with just a dataset of NBA players but quickly decided that there wasn't enough data. I was able to find datasets for NFL and MLB players. 
 
## NBA, NFL, and MLB raw Datasets

```{r warning=FALSE, message=FALSE}

library(tidyverse)
library(Lahman)
library(sf)
library(readr)
library(USAboundaries)
library(USAboundariesData)
library(maps)
library(leaflet)
library(knitr)


NBA <- read.csv("NBA_players.csv")

kable(NBA[1:4,], caption = "NBA")

NFL <- read.csv("nflprofiles.csv")
NFL <- arrange(NFL,college)
kable(NFL[1:4,], caption = "NFL")

MLB <- Lahman::CollegePlaying %>%
  left_join(Lahman::Schools, by = "schoolID") %>%
  left_join(Lahman::People, by = "playerID")
MLB <- arrange(MLB,name_full)
kable(MLB[1:4,], caption = "MLB")

```

## Cleaning the Data

I started with these three datasets. I had to rename certain columns on each to get down to the these 13 columns: FullName, College, BirthDay, DeathDate, BirthPlace, BirthCity, BirthState, Handedness, HighSchool, DraftTeam, DraftPick, DraftRound, DraftYear. 

The NFL dataset had different names, either the nicknames or the acronyms, from the MLB and NBA datasets. I went into the NFL Dataset and renamed each college to the correct full name that would align with the other datasets. 

After renaming those columns and selecting only them, I added a column in each df titled 'League' and in each df I either put 'MLB', 'NFL', or 'NBA'. Then I put them together using rbind.

I used a case_when() to add another column, 'Conference', that added the column for each school in the Power 5 conferences (ACC, BIG10, BIG12, PAC12, SEC). After some research and coding I figured who the top schools were and added 2 more columns for their Longitude and Latitude.

## Cleaned Dataframe

``` {r warning=FALSE, message=FALSE}

data <- read.csv("FinalReadyToCode.csv")

kable(data[1:4,],caption = "Final Dataset")


```

## Draft Overview

```{r warning=FALSE, message=FALSE}


Sport_Draft_counts <- count(data, League, DraftYear) %>%
  na.omit() %>%
  filter(
    DraftYear < 2012,
    DraftYear >1947,
    DraftYear != "1st"
  )

 Sport_Draft_counts$DraftYear <- as.numeric(Sport_Draft_counts$DraftYear)

ggplot(Sport_Draft_counts)+
  geom_point(mapping = aes(x = DraftYear, y = n, color = League)) +
  theme_bw()+
   theme(
    axis.text.x = element_text(angle = 75, hjust = 1)
  )+
    labs(x = "Year of Draft",
       y = "Total Number of Players Drafted",
       title = "Number of Players Drafted Each Year by Each League")+
  scale_x_continuous(breaks = seq(from = 1948, to = 2012, by = 2))



```

Looking at this there is a gradual obvious trend upward in the amount of draft picks each year. 1967 is a local high point for each and actually the max 3 of picks for each sport. The NBA and NFL go on to increase a little more and eventually even out starting after about 1978. 

## Breakdown of Colleges

```{r warning=FALSE, message=FALSE}



collegecounts <- count(data, College, Lat, Lon) %>%
  filter(
    College != "none",
    College != "na",
    College != "None"
  ) %>%
  na.omit()

ggplot(collegecounts %>%
         filter(
           n > 300
         )) +
  geom_col(mapping = aes(x = reorder(College,n), y = n), fill = "maroon") +
  coord_flip() +
  theme_bw()+
  labs(x = "College",
       y = "Count", 
       title = "Total Counts by College"
  )

```

You can see that Notre Dame and USC are far and away producing the most professional athletes. This data is skewed though, due to the large amount of football players in contrast to the other sports. There are 25,000 NFL players in this dataset while the NBA has about 5,000 and MLB has about 7,000.

```{r warning=FALSE, message=FALSE}
collegecountsbysport <- count(data, College, League) %>%
  filter(
    College != "none",
    College != "na",
    College != "None"
  ) %>%
  na.omit()


top50sport <- collegecountsbysport %>%
  group_by(League)%>%
  slice_max(order_by = n, n = 50) %>%
  mutate(rank = rank(-n))


ggplot(top50sport) +
  geom_col(mapping = aes(x = reorder(College,n), y = n, fill = League) ) +
  facet_wrap(~League,scales = "free") +
  theme_bw()+
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none")+
  labs(y = "Count",
       title = "Top 50 Colleges accross each League")

```

The top schools for MLB end up producing about 30 to 50.
The top schools for NFL end up producing about 170 to 300.
The top schools for NBA end up producing about 25 to 40.



## Percent of Leauge

```{r warning=FALSE, message=FALSE}


top10sport <- collegecountsbysport %>%
  group_by(League)%>%
  slice_max(order_by = n, n = 10) %>%
  mutate(rank = rank(-n))


## put image in

top10percent <- top10sport %>%
  mutate(
    percent_of_league = case_when(
      League == "MLB" ~ n/7550,
      League == "NFL" ~ n/25018,
      League == "NBA" ~ n/4685
    )
  )

top10percent <- arrange(top10percent, -percent_of_league)

kable(top10percent[1:10,], caption = "Percent of Leagues")

ggplot(top10percent) +
  geom_col(mapping = aes(x = College, y = percent_of_league, fill = League))+
  facet_wrap(~League) +
  coord_flip() +
  labs(x = "College",
       y = "Percent of League",
       title = "Percent of Each League by College")+
  theme_bw()+
  theme(
    legend.position = "none")


```

## Ranks

```{r warning=FALSE, message=FALSE}

collegecountsbysport <- count(data, College, League, Conference) %>%
  filter(
    College != "none",
    College != "na",
    College != "None"
  ) %>%
  na.omit()


topcollege <- collegecountsbysport %>%
  group_by(League)%>%
  slice_max(order_by = n, n = 1000) %>%
  mutate(rank = rank(-n)) %>%
  select(-n) %>%
  pivot_wider(names_from = League, values_from = rank) 

topcollege$finalrank <- topcollege$MLB+topcollege$NFL+topcollege$NBA

topcollegev2 <- topcollege %>%
  mutate(rank = rank(finalrank)) %>%
  filter(rank < 51)


finalrank <- arrange(topcollegev2,rank)
 

kable(finalrank[1:20,], caption = "Final Ranking System")

```

I came up with my own simple formula to determine the top school. I took the rank for each school in each sport and added it up and ranked the lowest total #1 and so on. 


## Breakdown by Conference

```{r warning=FALSE,message=FALSE}

dataConf <- data

dataConf$num <- 1

byConf <- dataConf %>%
  group_by(Conference, League) %>%
    summarise(
      Total = sum(num)
    ) %>%
  na.omit()


ggplot(data = byConf, aes(x = Conference, y = Total))+
  geom_col(aes(fill = League))+
  facet_wrap(~League, scales = "free") +
  theme_bw()+
  theme() +
  coord_flip()

```

The PAC12 is at the top in each league, leading me to believe it is the best producing conference.

```{r warning=FALSE, message=FALSE}
collegecountsbysport <- count(data, College, League, Lat, Lon, Conference) %>%
  filter(
    College != "none",
    College != "na",
    College != "None"
  ) %>%
  na.omit()


topcollege <- collegecountsbysport %>%
  group_by(League)%>%
  slice_max(order_by = n, n = 1000) %>%
  mutate(rank = rank(-n)) %>%
  select(-n) %>%
  pivot_wider(names_from = League, values_from = rank)

topcollege$finalrank <- topcollege$MLB+topcollege$NFL+topcollege$NBA

topcollegev2 <- topcollege %>%
  mutate(rank = rank(finalrank)) %>%
  filter(rank < 51)


top50 <- topcollegev2 %>%
  select(rank, College,Lat, Lon, Conference)

top50coordinates <- read_csv("top50.csv")

mapStates = map("state", fill = TRUE, plot = FALSE)

leaflet(data = mapStates) %>% addTiles() %>%
  addPolygons(fillColor = "white", stroke = FALSE) %>%
  addMarkers(data = top50,
             label = c(top50$rank,top50$College),
             labelOptions = labelOptions(noHide = T),
             clusterOptions = markerClusterOptions(),
             clusterId = "quakesCluster") 


```



```{r warning=FALSE,message=FALSE}


library(ggsflabel)
library(ggrepel)

top50coordinates <- read.csv("top50.csv")

top50v1 <- top50coordinates %>%
  mutate(geometry = gsub("c", "", geometry)) %>%
  mutate(geometry = gsub("\\(", "", geometry)) %>%
  mutate(geometry = gsub("\\)", "", geometry)) %>%
  mutate(geometry = str_trim(geometry)) %>%
  separate(geometry, c("lat", "long"), " ") %>%
  mutate(lat = as.numeric(lat),
         long = as.numeric(long))


states <- us_states()

ggplot()+
  geom_sf(data = states, fill = NA) +
  geom_point(data = top50v1, mapping = aes(x = lat, y = long, shape = Conference, color = rank), size = 1)+
  theme_bw()+
  guides(color = FALSE)+
  coord_sf(xlim = c(-125,-65), ylim = c(25,50))+
  labs(color = "Conference",
       title = "Breakdown of Top 50 by Conference")+
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )+
  geom_label_repel(data = top50v1, aes(x = lat, y = long, label = c(College)),size = 1, box.padding = 0.2)
  


```

The majority of the darker shades, which did indicate a better rank, are located on the eastern side of the country and in some cases clustered up. 

## Conclusion

After looking at all these I have come to conclude that the University of Michigan, UCLA, USC, and Notre Dame have been the top producers of professional athletes due to their high ranks and amount of athletes across each sport. 

## Future Plans

In the future I would like to try and find more data sets that have better data on High Schools attended as well as whole datasets for hockey, soccer, and women's sports. Finding and merging all these together can help me fully understand where the talent is best produced. I want to also look more at high schools and places of birth once I get more data to understand that better as well. I also want to find better Draft data so I can look at where the top talent is coming from historically. Salary and longevity are two other variables that I am interested in diving into to find significance there. Once I can get all these into 1 data set I can create more diverse graphs that look deeper into where athletes come from. 




For questions or comments, I can be reached at bradymonks@gmail.com