---
title: "Case Study 7"
author: "Brady Monks"
date: "10/13/2020"
output: 
  html_document:
    keep_md: TRUE
    code_folding: hide
---

## Loading Data In

```{r message=FALSE,warning=FALSE}

# Case Study 7

library(tidyverse)
library(haven)
library(readr)
library(downloader)
library(readxl)
library(foreign)

Tubheight <- tempfile()
download("https://github.com/WJC-Data-Science/DTS350/raw/master/Height.xlsx",
         Tubheight, mode = "wb")
Tubheight <- read_xlsx(Tubheight, skip = 2, col_names = TRUE)
Tubheight

germanheight <- read_dta("https://github.com/WJC-Data-Science/DTS350/raw/master/germanconscr.dta")
germanheight

germanprison <- read_dta("https://github.com/WJC-Data-Science/DTS350/raw/master/germanprison.dta")
germanprison


heightSE <- tempfile()
download("https://github.com/WJC-Data-Science/DTS350/raw/master/Heights_south-east/B6090.DBF", heightSE, mode = "wb")
heightSE <- read.dbf(heightSE)


heightscsv <- read_csv("https://raw.githubusercontent.com/hadley/r4ds/master/data/heights.csv")
heightscsv

heightsav <- tempfile()
download("http://www.ssc.wisc.edu/nsfh/wave3/NSFH3%20Apr%202005%20release/main05022005.sav", heightsav, mode = "wb")
heightsav <- read_sav(heightsav)
heightsav

householdsav <- tempfile()
download("http://www.ssc.wisc.edu/nsfh/wave3/NSFH3%20Apr%202005%20release/nsfh3mainroster1-04042005.sav", householdsav, mode = "wb")
householdsav <- read_sav(householdsav)
householdsav

```

Tubingen Data

``` {r message=FALSE,warning=FALSE}

Tubheighty <- Tubheight %>%
  pivot_longer(c(`1800`:`2011`), names_to = "year_decade", values_to = "height.cm") %>%
  na.omit() %>%
  separate(year_decade, into = c("century","decade"), sep = -2, remove = FALSE)%>%
  separate(decade, into = c("decade", "year"), sep = -1, remove = FALSE) %>%
  mutate(height.in = height.cm/2.54) %>%
  rename(country = "Continent, Region, Country") %>%
  select(Code, country, century, decade, year, height.cm, year_decade, height.in)

  
Tubheighty <- Tubheighty[order(Tubheighty$year_decade),]

write.csv(Tubheighty,"Tubheighty.csv")

```

## Cleaning Data Sets

``` {r message=FALSE,warning=FALSE}

clean_wisc_height <- heightsav %>%
          select(DOBY, RT216F, RT216I)%>%
          mutate(height.in = (RT216F*12)+ RT216I) %>%
          mutate(birth_year = DOBY + 1900) %>%
          select( -DOBY, -RT216F, -RT216I) %>%
          mutate(height.cm = height.in*2.54) %>%
          add_column(study = "Wisconsin") %>%
          select(birth_year,height.cm,height.in,study)
  
  
head(clean_wisc_height)


clean_BLS_height <- heightscsv %>%
              select(height) %>%
              add_column(birth_year = 1950) %>%
              add_column(study = "BLS") %>%
              mutate(height.in = height) %>%
              select(-height) %>%
              mutate(height.cm = height.in*2.54)%>%
  select(birth_year,height.cm,height.in,study)

head(clean_BLS_height)              

clean_SE_height <- heightSE %>%
          select(SJ, CMETER) %>%
          rename(birth_year = "SJ") %>%
          rename(height.cm = "CMETER") %>%
          mutate(height.in = height.cm/2.54)%>%
          add_column(study = "Southeast")%>%
  select(birth_year,height.cm,height.in,study)
  
head(clean_SE_height)


clean_German_height <- germanheight %>%
        select(bdec, height) %>%
        rename(birth_year = "bdec") %>%
        rename(height.cm = "height") %>%
        mutate(height.in = height.cm/2.54) %>%
        add_column(study = "German Males in Bavaria")%>%
  select(birth_year,height.cm,height.in,study)

head(clean_German_height)


clean_Prison_height <- germanprison %>%
  select(bdec, height) %>%
  rename(birth_year = "bdec") %>%
  rename(height.cm = "height") %>%
  mutate(height.in = height.cm/2.54) %>%
  add_column(study = "German Prison")%>%
  select(birth_year,height.cm,height.in,study)

head(clean_Prison_height)

#clean_Wisc_household <- householdsav %>%
#  select(doby, RT216F, RT216I)%>%
#  mutate(height.in = (RT216F*12)+ RT216I) %>%
#  mutate(birth_year = doby + 1900) %>%
#  select( -doby, -RT216F, -RT216I) %>%
#  mutate(height.cm = height.in*2.54) %>%
#  add_column(study = "Wisconsin")
 

new_data <- bind_rows(clean_BLS_height,clean_German_height,clean_Prison_height,clean_SE_height,clean_wisc_height)   
new_data    

write.csv(new_data,"new_data.csv")

```

## Plot

``` {r message=FALSE, warning=FALSE}

germany <- Tubheighty %>% 
  filter(`country` == "Germany")

ggplot() +
  geom_point(data = Tubheighty, mapping = aes(x = year_decade, y = height.cm), color = 'grey') +
  geom_point(data = germany, mapping = aes(x = year_decade, y = height.cm), color = 'maroon')+
  labs(x = "Birth Year",
       y = "Height (cm)",
       title = "Height in Germany")+
  theme_bw()


ggplot( data = new_data) +
  geom_point(mapping = aes(x = birth_year, y = height.cm, color = study)) +
  geom_smooth(mapping = aes(x = birth_year, y = height.cm))+
  labs(x = "Birth Year",
       y = "Height (cm)",
       title = "All Data")

wisc_data <- new_data %>% 
  filter(study == 'Wisconsin')
ggplot(data = wisc_data) +
  geom_point(mapping = aes(x = birth_year, y = height.cm, color = height.in)) +
  geom_smooth(mapping = aes(x = birth_year, y = height.cm)) +
  theme_bw()+
  labs(x = "Birth Year",
       y = "Height (cm)",
       title = "Wisconsin Data")



```

 ## Summary
 
 From the Tubingen data you can see there is a slight decline in height from 1810 to 1850, however since then there has been a steady incline to this day. From the all data graph, you can see there isn't much variation throughout history. It generally hovers around 170ish. The Wisconsin data set had the biggest range of data, so I looked at that and, as like the others, there was little to no variation whatsoever. 

